Article Title

### 2013-12-30 plos.tex
minor

\begin{document}
% Title must be 150 characters or less
\begin{flushleft}
{\Large
Improved estimation of neural correlations suggests detailed interactions in visual cortex
}
% Insert Author names, affiliations and corresponding author email.
\\
Dimitri Yatsenko,$^{1,\ast}$, 
Kre\v{s}imir Josi\'{c}$^{2}$,
Alexander S.~Ecker$^{1,3,4}$,
Emmanouil Froudarakis$^{1}$,
R.~James Cotton$^{1}$,
Andreas S.~Tolias$^{1,5}$
\\
\bf{1} Department of Neuroscience, Baylor College of Medicine, Houston, TX, USA
\\
\bf{2} Department of Mathematics and Department of Biology and Biochemistry, University of Houston, Houston, TX, USA
\\
\bf{3}  Werner Reichardt Center for Integrative Neuroscience and Institute for Theoretical Physics, University of T\"ubingen, Germany
\\
\bf{4} Bernstein Center for Computational Neuroscience, T\"ubingen, Germany
\\
\bf{5} Department of Computational and Applied Mathematics, Rice University, Houston, TX, USA

@@

In this study, we sought most statistically efficient estimators of neural correlation matrices in recordings from large, dense groups of cortical neurons.  Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 200 $\mu$m wide and 100 $\mu$m deep (150--350 cells) in mouse visual cortex.  We hypothesized that in these dense recordings, the correlation matrix would be most efficiently represented by a sparse network of partial correlations between pairs of observed neurons combined with common latent factors representing unobserved inputs and global network fluctuations.  Indeed, in cross-validation tests, the covariance matrix estimator based on such approximations outperformed other regularized estimators. Mirroring previously described patterns of synaptic connectivity, the density of positive partial correlations decreased rapidly with both physical distances and preferred orientation differences whereas negative partial correlations were less selective.  The inferred functional connectivity obtained using these methods can generate hypotheses about synaptic architecture that can guide future experimental work.

@@

For example, one form of low-dimensional representation of the correlation matrix is its low-rank approximation. Such approximations are particularly suitable for capturing shared fluctuations across the recorded population, which may arise in local circuits or be externally driven. Low-rank components can be extracted using principal component or factor analysis. The resulting structure of neural correlations has been analyzed in a number of studies with particular attention paid to their temporal dynamics \cite{Yu:2009}.

Alternatively, the correlation matrix can be expressed through the partial pairwise correlations. The partial pairwise correlation is the linear correlation remaining after accounting for correlations with all the other neurons. Partial correlations carry particular significance when the relationship between the variables is approximately linear. In this case, small pairwise partial correlations indicate that the two variables are nearly conditionally independent.

@@

\paragraph{Illustration of estimation of a neural correlation matrix}
To illustrate the challenges in estimating the correlation matrix from a finite sample, we consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972}. Here the estimate is produced by fitting only an optimal subset of the coefficients of the precision matrix while setting the rest to zero. We estimated the covariance matrix of somatic calcium signals of a group of 298 neurons in mouse visual cortex using both the sample covariance matrix (\figref{01}{A, B}) and covariance selection (\figref{01}{C}). Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are higher than typical firing rate correlations on shorter temporal scales. The close similarity of the two estimates of the correlation matrix (\figref{01}{B and C}) belies the utterly different partial correlation structure (\figref{01}{D and E}): The regularized estimate is produced from the same data by setting to zero 31,501 (71.2\%) of the off-diagonal coefficients of the precision matrix and fitting the remaining 12,752 coefficients. Estimation by covariance selection is attractive for several reasons: First, with fewer free parameters it is less susceptible to sampling noise, yielding estimates that are, on average, closer to the truth. Additionally, if we assume that partial correlations do indeed reflect elements of the functional connectivity in the circuit, the second estimate also appears more plausible: in neural circuits, each neuron only interacts with a fraction of the entire population.

\begin{figure}[!ht]    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
    {\caption{{\bf Illustration of regularized estimation of partial correlations.}
        {\bf A}. The sample correlation matrix of unprocessed somatic calcium signals from a population of cells in mouse visual cortex.
        The outlined square fragment is magnified in {\bf B}.
        {\bf C}. The same fragment of another estimate of the correlation matrix regularized to yield sparse partial correlations.
        Corresponding fragments of partial correlations matrices of the unregularized and regularized estimated are shown in {\bf D} and {\bf E}, respectively.
    }
    \label{fig:01}}
    {\includegraphics[width=8.3cm]{./figures/Figure01.pdf}}
\end{figure}
\paragraph{Model selection}
Covariance selection is but one of many conceivable low-dimensional approximations of the covariance matrix. Indeed any probabilistic model of the joint population activity regularized toward a small number of parameters automatically defines a low-dimensional approximation of the covariance matrix. Models that accurately reflect the dominant regularities in the activity of the circuit, its so-called `functional connectivity', tend to produce estimates that are nearest to the true value of the covariance matrix and are said to be more \emph{efficient} or to \emph{dominate} other estimators. An estimator's superior efficiency can be ascertained by cross-validation.  Because different types of neural circuits have different types of regularities in their activity, different probabilistic models and covariance matrix estimators may dominate in each case. The estimator that is shown to dominate all others for a specific type of neural circuit serves as an effective descriptor the circuit’s correlation structure.  This structure may then be analyzed with respect to the circuit's anatomical organization.
\paragraph{Summary of findings}
In this study, we compared four regularized covariance matrix estimators biased toward different respective low-dimensional correlation structures: `shrinkage toward diagonal' or $C_{\sf diag}$, `shrinkage toward a multifactor model' or $C_{\sf factor}$, `sparse partial correlations' or $C_{\sf sparse}$, and `sparse partial correlations with latent units' or $C_{\sf sparse+latent}$.  First, we demonstrated that, in simulations with known true low-dimensional correlation structures, regularized estimators with matching types of low-dimensional structure generally dominated over the other estimators. We then used cross-validation to determine the most efficient estimator for the population activity of dense groups of neurons in mouse primary visual cortex. We found that $C_{\sf sparse+latent}$ consistently dominated the other estimators.  Estimate produced by $C_{\sf sparse+latent}$ revealed networks of interactions that differed substantially from networks of strongest correlations and depended strongly on the physical distance separating pairs of cells and on the differences in their preferred orientations.

\section*{Results}
% Results and Discussion can be combined.
\paragraph{Covariance estimation}
We considered four regularized estimators based on distinct families of low-dimensional target estimates: `independent', `latent factors', `sparse interactions', and `sparse+latent' (\figref{02}{row 1}).

@@

The target of the second regularized estimator $C_{\sf factor}$, is the factor model $LL^\T + \Psi$, where $L$ is the $p\times d$ matrix of \emph{factor loadings}, where $d$ is the number of factors, and $\Psi$ is the diagonal matrix of independent variances. The estimator, 
\begin{equation}\label{eq:c-factor}
C_{\sf factor} = (1-\lambda) C_{\sf 0} + \lambda (LL^\T + \Psi),
\end{equation}
has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$. The target estimate $LL^\T + \Psi$ has the structure that arises when correlated fluctuations in population activity are linearly driven by a small number of latent factors that affect many cells while direct interactions between cells are insignificant (\figref{02}{B}).

@@

The fourth estimator $C_{\sf sparse+latent}$, combines the assumptions of the previous two.  The underlying structure arises from sparse interactions between the recorded neurons and latent factors driving their activity (\figref{02}{D}). The estimator has the form \cite{Chandrasekaran:2010,Ma:2013}:
\begin{equation}\label{eq:c-sl}
C_{\sf sparse+latent} = (S - LL^\T)^{-1}
\end{equation}
where, as above, $S$ is a sparse matrix and $L$ is a $d\times p$ matrix of factor loadings. The estimator has two hyperparameters: the number of latent units $d$ and the sparsity of $S$.

@@

When the ground truth, $\Sigma,$ is not accessible, loss can be estimated solely from the data through \emph{validation}.  In validation, an additional independent \emph{testing sample} is used to compute a sample covariance estimate $C_{\sf 0}^\prime$.  Then \emph{validation loss} $\loss{C,C_{\sf 0}^\prime}$ can be used as an unbiased estimate\footnote
{
    Validation loss $\loss{C,C_{\sf 0}^\prime}$ is an unbiased estimate of loss $\loss{C,\Sigma}$ when $\loss{\cdot,\cdot}$ is \emph{additive} in its second argument so that 
 \begin{equation*}\label{eq:additivity}
 \loss{C,z_1} + \loss{C,z_2} \equiv \loss{C,z_1+z_2}
 \end{equation*}

Then the absence of bias can be shown:
\begin{equation*}
    \E[C]{\E[C_{\sf 0}^\prime]{\loss{C,C_{\sf 0}^\prime}}}=\E[C]{\loss{C,\E[C_{\sf 0}]{C_{\sf 0}^\prime}}}=\E[C]{\loss{C,\Sigma}}
\end{equation*}

The last result expression is known as the \emph{estimator risk} which serv

The property of additivity does not hold for other popular loss functions such as Stein's entropy loss or various quadratic losses, for example, which prevents their substitution with corresponding validation losses. 
} 
of loss $\loss{C,\Sigma}$.  Thus, estimators resulting in consistently lower validation loss can be inferred to produce estimates that are closer to truth than estimators with higher validation loss.

@@

In these highly localized populations both direct interactions between cells and common diffuse inputs likely contribute to overall population co-variability.  We therefore hypothesized that regularized estimates capturing a sparse partial correlation structure, and/or interactions with latent units would be the most efficient. At the same time, average correlations were relatively small (\figref{03}{G}), giving a potential advantage to estimates biased toward independence (\emph{e.g.}\;$C_{\sf diag}$). Finally, correlation structure could possibly be best explained by common fluctuations across the entire population, to the advantage of estimates biased toward low-rank correlation structure (\emph{e.g.}\;$C_{\sf factor}$, $C_{\sf factor+lowrank}$). Thus it was \emph{a priori} unclear which estimator would perform best.

@@

\begin{figure}[!ht]    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
    {\caption{{\bf The sparse+latent estimator $C_{\sf sparse+latent}$ outperforms the other estimators on neural data.}
    {\bf A--D.} Histograms of average cross-validation loss differences of the respective estimators $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ from $C_{\sf sparse+latent}$. 
    The histograms are based on 31 imaged sites in 24 mice. 
    All medians (red dashed lines) were significantly greater than zero, indicating the dominance of $C_{\sf sparse+latent}$ over the other estimators. 
    The green arrows indicate the results for the site shown in Fig.~\ref{fig:03} and Fig.~\ref{fig:05}
    } \label{fig:04}}
    {\includegraphics{./figures/Figure04.pdf}}
\end{figure}

@@

Another reduced representation of correlation matrices is obtained by sparsifying its inverse, resulting in a graph of partial correlations. Under the assumption of a multivariate normal distribution or, more generally, an elliptical distribution, sparse partial correlations suggest a sparse network of dependencies, \emph{i.e.}\;a probabilistic graphical model. This property has led some investigators to propose that sparse inverse approximation of correlations may serve as closer estimates of direct anatomical interactions between brain regions in imaging studies than correlations \cite{Varoquaux:2012,Ryali:2012}. However, since deviations from elliptical distribution disrupt the correspondence between partial correlations and linear dependencies \cite{Loh:2012}, it is not clear whether this is indeed  the case. Partial correlations are meaningful when most interacting variables are observed. Perhaps for this reason, until now, partial correlations had not been considered in analysis of multineuronal recordings.

@@

The correlation structures we considered also corresponded to structures of conditional dependencies in multivariate Gaussian models or, more generally, elliptical distributions. For example, zero partial correlation indicates conditional independence between a pair of variables under an elliptical distribution, but this correspondence weakens or breaks down with deviation from elliptical distributions \cite{Loh:2012}.  Other probabilistic models, when optimally sparsified for best cross-validated estimation of the covariance matrix, could suggest very different forms of low-dimensional representations (structures) of the covariance matrix. For example, spatiotemporal Ising models, marginalized to reproduce the correlation matrix at the same temporal scale, would suggest an alternative correlation structure from the same data. At this time we do not have computationally efficient methods for finding sparse structures of dependencies in Ising models optimized with respect to cross-validated estimation of their cross-correlation matrices. If we did, and the estimator thus constructed outperformed our $C_{\sf sparse+lowrank}$ estimator, we would be adopt this new estimator as one producing the best structure of the connectivity. Various forms of GLMs could also be evaluated with in the same way: sparsified for best estimation of the covariance matrix (not log-likelihood) and compared to the estimators studied in this study.  Again, if this new covariance estimator outperforms, $C_{\sf sparse+lowrank}$, we will be compelled to adapt it for estimating the functional connectivity.

@@

Estimator $C_{\sf factor}$ used two hyperparameters: the number of latent factors $d$ and the shrinkage intensity $\lambda \in [0, 1]$.  
The $p\times d$ factor loading matrix $L$ and individual variances $\Psi$ were computed by solving the minimization problem
\begin{equation}
(L,\Psi) = \argmin\limits_{\tilde L,\tilde\Psi} \loss{\tilde L \tilde L^\T + \tilde\Psi,C_{\sf 0}},
\end{equation}
which we solved by an expectation-maximization (EM) algorithm.  Under our chosen loss function (Eq.~\ref{eq:loss}), this is equivalent to maximum likelihood estimation of $L$ and $\Psi$ under the multivariate Gaussian distribution.  The final regularized estimate is obtained by linear shrinkage toward the factor model (Eq.~\ref{eq:c-factor}).

@@

Estimator $C_{\sf sparse+lowrank}$ has two hyperparameters: the number of latent units $d$ and the sparsity level $\lambda$. It estimates the larger sparse precision matrix $S^\ast$ of the joint distribution of the $p$ observed neurons and $d$ latent units.  
\begin{equation}
S^\ast=
\begin{pmatrix}
S & S_{12} \\
S_{12}^\T & S_{22}
\end{pmatrix},
\end{equation}
where the $p\times p$ partition $S$ corresponds to the visible units and expresses their partial correlation structure, and $S_{12}$ and $S_{22}$ are of size $p\times d$ and $d\times d$, respectively.
Then the covariance matrix of the observed population is 
\begin{equation}
C_{\sf sparse+lowrank} = (S^\ast)^{-1} = \left(S-S_{12}S_{22}^{-1}S_{12}^\T\right)^{-1}
\end{equation}
The matrix $ S_{12}S_{22}^{-1}S_{12}^\T$ has rank $d$. Rather than searching for the optimal sparse structure of $S_{12}$ and $S_{22}$, an ill-posed problem, we estimated these components together as a single matrix $LL^\T$ of rank $d$, where $L$ is of size $p\times d$.
The estimate is found in two steps. First, we use the ADMM algorithm to find the zero structure of $S$ by minimizing the $L_1$-penalized loss \cite{Chandrasekaran:2010,Ma:2013}:
\begin{equation}
(Z,\cdot) = \argmin\limits_{\tilde Z,\tilde L} \loss{\tilde Z-\tilde L\tilde L^\T} + \|\tilde Z\|_1
\end{equation}
Then we find the sparse and low-rank components of the precision matrix by minimizing the loss 
\begin{equation}
(S,L) = \argmin\limits_{\tilde S \in Z^\sharp,\tilde L} \loss{\tilde S-\tilde L \tilde L^\T}
\end{equation}

The partial correlation matrix computed from the overall estimate, computed from $C_{\sf sparse+lowrank}$ according to Eq.~\ref{eq:partial}, includes the effects of interactions between the visible and latent units.  This estimate of the partial correlations was used in analyses where sparsity was not required (\figref{05}{B and E} and \figref{06}{A}, and \figref{07}{A--C}).  For analyses that required a sparse network (\figref{05}{F, G, H} and \figref{06}{B, D} and \figref{07}{D--F}) was computed using the extended precision matrix $S^\ast$:
\begin{equation}
P_{\sf sparse} = (S\circ I)^{-\frac 1 2} S  (S\circ I)^{-\frac 1 2}
\end{equation}

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics{./figures/Figure-Supp02.pdf}
\end{center}
\caption{{\bf All-to-all performance comparisons of the sample covariance matrix and the four regularized estimators with respect to quadratic cross-validation loss.}
}
\label{supp:02}
\end{figure}

>>>>

\begin{document}
% Title must be 150 characters or less
\begin{flushleft}
{\Large
Inferred partial neural correlations suggest detailed interactions in cortical circuits
}
% Insert Author names, affiliations and corresponding author email.
\\
Dimitri Yatsenko,$^{1,\ast}$, 
Kre\v{s}imir Josi\'{c}$^{2}$,
Alexander S.~Ecker$^{1,3,4}$,
Emmanouil Froudarakis$^{1}$,
R.~James Cotton$^{1}$,
Andreas S.~Tolias$^{1,5}$
\\
\bf{1} Department of Neuroscience, Baylor College of Medicine, Houston, TX, USA
\\
\bf{2} Department of Mathematics and Department of Biology and Biochemistry, University of Houston, Houston, TX, USA
\\
\bf{3}  Werner Reichardt Center for Integrative Neuroscience and Institute for Theoretical Physics, University of T\"ubingen, Germany
\\
\bf{4} Bernstein Center for Computational Neuroscience, T\"ubingen, Germany
\\
\bf{5} Department of Computational and Applied Mathematics, Rice University, Houston, TX, USA

@@

In this study, we sought most statistically efficient estimators of neural correlation matrices in recordings from large, dense groups of cortical neurons.  Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 200 $\mu$m wide and 100 $\mu$m deep (150--350 cells) in mouse visual cortex.  We hypothesized that in these dense recordings, the correlation matrix would be most efficiently represented by a sparse network of partial correlations between pairs of observed neurons combined with common latent factors representing unobserved inputs and global network fluctuations.  Indeed, in cross-validation tests, the covariance matrix estimator based on such approximations outperformed other regularized estimators. Mirroring previously described patterns of synaptic connectivity, the density of positive partial correlations decreased rapidly with both physical distances and preferred orientation differences whereas negative partial correlations were less selective.  Because of its superior estimation performance, the inferred partial correlation structure obtained using these methods may provide a better description of the functional connectivity in dense circuits than the usual sample correlations. 

@@

For example, one form of low-dimensional representation of the correlation matrix is its low-rank approximation. Such approximations are particularly suitable for capturing shared fluctuations across the recorded population, which may arise in local circuits or be externally driven. Low-rank components can be extracted using principal component or factor analysis.

Alternatively, the correlation matrix can be expressed through the partial pairwise correlations. The partial correlation between a pair of neurons is a measure of their linear association after the linear effects of all other neurons have been taken into account. Partial correlations carry particular significance when the relationship between the variables is approximately linear. In this case, small pairwise partial correlations indicate that the two variables are nearly independent when conditioned on the activity of the other neurons.

@@

\paragraph{Illustration of estimation of a neural correlation matrix}
To illustrate the challenges and techniques of estimation of the correlation matrix from a finite sample, we consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972}. Here the estimate is produced by fitting only an optimal subset of the coefficients of the precision matrix while setting the rest to zero. We estimated the covariance matrix of somatic calcium signals of a group of 298 neurons in mouse visual cortex using both the sample covariance matrix (\figref{01}{A, B}) and covariance selection (\figref{01}{C}). Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are higher than typical firing rate correlations on shorter temporal scales. The close similarity of the two estimates of the correlation matrix (\figref{01}{B and C}) belies the utterly different partial correlation structure (\figref{01}{D and E}): The regularized estimate is produced from the same data by setting to zero 31,501 (71.2\%) of the off-diagonal coefficients of the precision matrix and fitting the remaining 12,752 coefficients. Estimation by covariance selection is attractive for several reasons: First, with fewer free parameters it is less susceptible to sampling noise, yielding estimates that are, on average, closer to the truth. Additionally, if we assume that partial correlations do indeed reflect elements of the functional connectivity in the circuit, the second estimate also appears more plausible: in neural circuits, each neuron only interacts with a fraction of the entire population.

\begin{figure}[!ht]    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
    {\caption{{\bf Illustration of regularized estimation of partial correlations.}
        {\bf A}. The sample correlation matrix of unprocessed somatic calcium signals from a population of cells in mouse visual cortex.
        The outlined square fragment is magnified in {\bf B}.
        {\bf C}. The same fragment of another estimate of the correlation matrix regularized to yield sparse partial correlations.
        Corresponding fragments of partial correlations matrices of the unregularized and regularized estimated are shown in {\bf D} and {\bf E}, respectively.
    }
    \label{fig:01}}
    {\includegraphics[width=8.3cm]{./figures/Figure01.pdf}}
\end{figure}
\paragraph{Our contribution}
Various other regularization schemes also produce improved estimates.  In this study, we compared the sample covariance matrix and four regularized estimators based on distinct correlation structures as follows:
\begin{description}
\item[$C_{\sf 0}$] -- sample covariance matrix, the usual estimator used in previous studies.
\item[$C_{\sf diag}$] -- linear shrinkage toward a diagonal covariance matrix.
\item[$C_{\sf factor}$] -- linear shrinkage toward a low-rank matrix produced by factor analysis with an optimal number of factors.
\item[$C_{\sf sparse}$] -- sparse inverse \emph{covariance selection} \cite{Dempster:1972, Friedman:2008}, without shrinkage.   
\item[$C_{\sf sparse+latent}$] -- inverse matrix as the sum of a sparse  and a low-rank component \cite{Chandrasekaran:2010, Ma:2013}.
\end{description}

First, we demonstrated that, in simulations with known correlation structures, regularized estimators with matching structure were generally more efficient than the other estimators.

We also performed a cross-validated evaluation to establish the most efficient estimator for the population activity of dense groups of neurons in mouse primary visual cortex. We found that $C_{\sf sparse+latent}$ consistently outperformed the other estimators.  This estimator produced a sparse network of partial correlations between the observed neurons combined with correlations between several inferred latent units. Since it led to better estimates, we proposed that this structure provided a more physiologically relevant representation of the circuit’s functional connectivity than that obtained from the sample correlation matrix.  We showed that the partial correlation structure had stronger dependence on physical distances and differences in orientation preference than sample correlations. In Discussion, we outline possibly approaches to corroborate the relationship between the structure of the functional connectivity revealed by this and similar method to the anatomical organization of the circuit.

\section*{Results}
% Results and Discussion can be combined.
\paragraph{Covariance estimation}
We considered four regularized estimators based on distinct families of low-dimensional target estimates: $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$. In probabilistic systems with linear dependencies, these correspond to graphical modes with distinct types of structure  (\figref{02}{row 1}).

@@

The target of the second regularized estimator $C_{\sf factor}$, is the factor model $L + \Psi$, where $L$ is the $p\times p$ positive definite matrix of rank $d$, and $\Psi$ is the diagonal matrix of independent variances.  The estimator, 
\begin{equation}\label{eq:c-factor}
C_{\sf factor} = (1-\lambda) C_{\sf 0} + \lambda (L + \Psi),
\end{equation}
has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$. The target estimate $L + \Psi$ has the structure that arises when correlated fluctuations in population activity are linearly driven by a small number of latent factors that affect many cells while direct interactions between cells are insignificant (\figref{02}{B}).

@@

The fourth estimator, $C_{\sf sparse+latent}$, combines the assumptions of the previous two.  The underlying structure arises from sparse interactions between the observed neurons and several latent units (\figref{02}{D}). The estimator has the form \cite{Chandrasekaran:2010,Ma:2013}:
\begin{equation}\label{eq:c-sl}
C_{\sf sparse+latent} = (S - L)^{-1}
\end{equation}
where, as above, $S$ is a sparse matrix and $L$ is a $p\times p$ matrix of rank $d$. The estimator has two hyperparameters: the number of latent units $d$ and the sparsity of $S$.

@@

When the ground truth, $\Sigma$, is not accessible, loss can be estimated solely from the data through \emph{validation}.  In validation, an additional independent \emph{testing sample} is used to compute a sample covariance estimate $C_{\sf 0}^\prime$.  Then \emph{validation loss} $\loss{C,C_{\sf 0}^\prime}$ can be used as an estimate of loss $\loss{C,\Sigma}$.  Thus, estimators resulting in consistently lower validation loss can be inferred to produce estimates that are closer to truth than estimators with higher validation loss.

The negative log-likelihood loss (Eq.~{\ref{eq:loss}) is particularly convenient because it is additive in its second argument in the sense that 
 \begin{equation*}\label{eq:additivity}
 \loss{C,z_1} + \loss{C,z_2} \equiv \loss{C,z_1+z_2}
 \end{equation*}

Then the cross-validation loss is in fact an unbiased estimate of the regular loss.
\begin{equation*}
    \E[C]{\E[C_{\sf 0}^\prime]{\loss{C,C_{\sf 0}^\prime}}}=\E[C]{\loss{C,\E[C_{\sf 0}]{C_{\sf 0}^\prime}}}=\E[C]{\loss{C,\Sigma}}
\end{equation*}

The property of additivity does not hold for other popular loss functions such as Stein's entropy loss or various quadratic losses.  With non-additive loss functions, cross-validation loss is a biased estimate of loss, but it is still generally appropriate to be used in lieu of loss.

@@

In these highly localized populations both direct interactions between cells and common diffuse inputs likely contribute to overall population co-variability.  We therefore hypothesized that regularized estimates capturing a sparse partial correlation structure, and/or interactions with latent units would be the most efficient. At the same time, average correlations were relatively small (\figref{03}{G}), giving a potential advantage to estimates biased toward independence (\emph{e.g.}\;$C_{\sf diag}$). Finally, correlation structure could possibly be best explained by common fluctuations across the entire population, to the advantage of estimates biased toward low-rank correlation structure (\emph{e.g.}\;$C_{\sf factor}$, $C_{\sf sparse+latent}$). Thus it was \emph{a priori} unclear which estimator would perform best.

@@

\begin{figure}[!ht]    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
    {\caption{{\bf The sparse+latent estimator $C_{\sf sparse+latent}$ outperforms the other estimators on neural data.}
    {\bf A--D.} Histograms of average cross-validation loss differences of the respective estimators $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ from $C_{\sf sparse+latent}$. 
    The histograms are based on 31 imaged sites in 24 mice. 
    All medians (red dashed lines) were significantly greater than zero, indicating the dominance of $C_{\sf sparse+latent}$ over the other estimators. 
    The arrow heads indicate the results for the site shown in Fig.~\ref{fig:03} and Fig.~\ref{fig:05}
{\bf A} $C_{\sf sparse+latent}$ outperforms $C_{\sf 0}$: median improvement 0.16 nats/neuron, $p=3.9\times 10^{-5}$.
    {\bf B} $C_{\sf sparse+latent}$ outperforms $C_{\sf diag}$: median improvement 0.0032 nats/neuron, $p=2.2\times 10^{-3}$.
    {\bf C} $C_{\sf sparse+latent}$ outperforms $C_{\sf factor}$: median improvement 0.16 nats/neuron, $p=6.9\times 10^{-5}$.
    {\bf D} $C_{\sf sparse+latent}$ outperforms $C_{\sf sparse}$: median improvement 0.0016 nats/neuron, $p=5.5\times 10^{-6}$.
} \label{fig:04}}
    {\includegraphics{./figures/Figure04.pdf}}
\end{figure}

@@

Another reduced representation of correlation matrices is obtained by sparsifying its inverse, resulting in a graph of partial correlations. Under the assumption of a multivariate normal distribution or, more generally, an elliptical distribution, sparse partial correlations suggest a sparse network of dependencies, \emph{i.e.}\;a probabilistic graphical model. This property has led some investigators to propose that sparse inverse approximation of correlations may serve as closer estimates of direct anatomical interactions between brain regions in imaging studies than correlations \cite{Varoquaux:2012, Ryali:2012}. However, since deviations from elliptical distribution disrupt the correspondence between partial correlations and linear dependencies \cite{Loh:2012}, it is not clear whether this is indeed  the case. Partial correlations are meaningful when most interacting variables are observed. Perhaps for this reason, until now, partial correlations had not been considered in analysis of multineuronal recordings.

@@

The correlation structures we considered also corresponded to structures of conditional dependencies in multivariate Gaussian models or, more generally, elliptical distributions. For example, zero partial correlation indicates conditional independence between a pair of variables under an elliptical distribution, but this correspondence weakens or breaks down with deviation from elliptical distributions \cite{Loh:2012}.  Other probabilistic models, when optimally sparsified for best cross-validated estimation of the covariance matrix, could suggest very different forms of low-dimensional representations (structures) of the covariance matrix. For example, spatiotemporal Ising models, marginalized to reproduce the correlation matrix at the same temporal scale, would suggest an alternative correlation structure from the same data. At this time we do not have computationally efficient methods for finding sparse structures of dependencies in Ising models optimized with respect to cross-validated estimation of their cross-correlation matrices. If we did, and the estimator thus constructed outperformed our $C_{\sf sparse+latent}$ estimator, we would be adopt this new estimator as one producing the best structure of the connectivity. Various forms of GLMs could also be evaluated with in the same way: sparsified for best estimation of the covariance matrix (not log-likelihood) and compared to the estimators studied in this study.  Again, if this new covariance estimator outperforms, $C_{\sf sparse+latent}$, we will be compelled to adapt it for estimating the functional connectivity.

@@

Estimator $C_{\sf factor}$ used two hyperparameters: the number of latent factors $d$ and the shrinkage intensity $\lambda \in [0, 1]$.  
The $p\times d$ factor loading matrix $L$ and individual variances $\Psi$ were computed by solving the minimization problem
\begin{equation}
(L,\Psi) = \argmin\limits_{\tilde L,\tilde\Psi} \loss{\tilde L + \tilde\Psi,C_{\sf 0}},
\end{equation}
which we solved by an expectation-maximization (EM) algorithm.  Under our chosen loss function (Eq.~\ref{eq:loss}), this is equivalent to maximum likelihood estimation of $L$ and $\Psi$ under the multivariate Gaussian distribution.  The final regularized estimate is obtained by linear shrinkage toward the factor model (Eq.~\ref{eq:c-factor}).

@@

Estimator $C_{\sf sparse+latent}$ has two hyperparameters: the number of latent units $d$ and the sparsity level $\lambda$. It estimates the larger sparse precision matrix $S^\ast$ of the joint distribution of the $p$ observed neurons and $d$ latent units.  
\begin{equation}
S^\ast=
\begin{pmatrix}
S & S_{12} \\
S_{12}^\T & S_{22}
\end{pmatrix},
\end{equation}
where the $p\times p$ partition $S$ corresponds to the visible units and expresses their partial correlation structure, and $S_{12}$ and $S_{22}$ are of size $p\times d$ and $d\times d$, respectively.
Then the covariance matrix of the observed population is 
\begin{equation}
C_{\sf sparse+latent} = (S^\ast)^{-1} = \left(S-S_{12}S_{22}^{-1}S_{12}^\T\right)^{-1}
\end{equation}
The matrix $ S_{12}S_{22}^{-1}S_{12}^\T$ has rank $d$. Rather than searching for the optimal sparse structure of $S_{12}$ and $S_{22}$, an ill-posed problem, we estimated these components together as one positive-definite $p\times p$ matrix $L$ of rank $d$.

The estimate is found in two steps. First, we use the ADMM algorithm to find the zero structure of $S$ by minimizing the $L_1$-penalized loss \cite{Chandrasekaran:2010,Ma:2013}:
\begin{equation}
(Z,\cdot) = \argmin\limits_{\tilde Z,\tilde L} \loss{\tilde Z-\tilde L} + \|\tilde Z\|_1
\end{equation}
Then we find the sparse and low-rank components of the precision matrix by minimizing the loss 
\begin{equation}
(S,L) = \argmin\limits_{\tilde S \in Z^\sharp,\tilde L} \loss{\tilde S-\tilde L }
\end{equation}

The partial correlation matrix computed from the overall estimate, computed from $C_{\sf sparse+latent}$ according to Eq.~\ref{eq:partial}, includes the effects of interactions between the visible and latent units.  This estimate of the partial correlations was used in analyses where sparsity was not required (\figref{05}{B and E} and \figref{06}{A}, and \figref{07}{A--C}).  For analyses that required a sparse network (\figref{05}{F, G, H} and \figref{06}{B, D} and \figref{07}{D--F}) was computed using the extended precision matrix $S^\ast$:
\begin{equation}
P_{\sf sparse} = (S\circ I)^{-\frac 1 2} S  (S\circ I)^{-\frac 1 2}
\end{equation}

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics{./figures/Figure-Supp02.pdf}
\end{center}
\caption{{\bf The sparse+latent estimator outperforms the other estimators with respect to quadratic cross-validation loss $\loss{C,C_{\sf 0}^\prime}=\frac 1 {p^2}\Tr(C^{-1}C_{\sf 0}^\prime-I)^2$.}
}
\label{supp:02}
\end{figure}
